How Optimal is Word-Referent Identification Under Multimodal Uncertainty?

### 2018-05-29 Writeup/ms.tex
last pass

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Word-Referent Identification Under Multimodal Uncertainty},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{Word-Referent Identification Under Multimodal Uncertainty}

@@

  \abstract{Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We introduce
a task that allows us to examine how adults identify words under joint
uncertainty in the auditory and visual modalities. We propose an ideal
observer model which provides an account of how auditory and visual cues
are combined optimally. Model predictions are tested in three
experiments where word recognition is made under two kinds of
uncertainty: category ambiguity and/or distorting noise. In all cases,
the optimal model explains much of the variance in human mean judgments.
In particular, when the signal is not distorted with noise, participants
weight the auditory and visual cues optimally, that is, according to the
relative reliability of each modality. But when one modality has noise
added to it, human perceivers systematically prefer the unperturbed
modality to a greater extent than the optimal model does. The study
provides a formal framework which helps us understand precisely how word
form and word meaning interact in word recognition under uncertainty.
Moreover it offers a first step towards a model that accounts for
form-meaning synergy in early word learning.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

    \indent Word count: X
  }

@@

Language uses symbols expressed in one modality, e.g., the auditory
modality, in the case of speech, to communicate about the world, which
we perceive through many different sensory modalities. Consider hearing
someone yell \enquote{bee!} at a picnic, as a honey bee buzzes around
the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence. However, word identification
takes place in a noisy world, and the cues received through each
modality may not provide a definitive answer. On the auditory side,
individual acoustic word tokens are almost always ambiguous with respect
to the particular sequence of phonemes they represent, which is due to
the inherent variability of how a phonetic category is realized
acoustically (Hillenbrand, Getty, Clark, \& Wheeler, 1995). And some
tokens may be distorted additionally by mispronunciation or ambient
noise. Perhaps the speaker was yelling \enquote{pea} and not
\enquote{bee}. Similarly, a sensory impression may not be enough to make
a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

The task of matching the sound to the corresponding visual object has
been extensively studied in the developmental literature since it is
considered to be an crucial instance of early word learning. For
example, many studies focused on how children might succeed in this task
despite high referential ambiguity (Medina, Snedeker, Trueswell, \&
Gleitman, 2011; Pinker, 1989; Smith \& Yu, 2008; Suanda, Mugwanya, \&
Namy, 2014; Vlach \& Johnson, 2013; Vouloumanos, 2008; Yurovsky \&
Frank, 2015). However, even when they know the exact meanings of the
words, listeners (both children and adults) often face the task of
recognizing which word the speaker utters, especially under noisy
circumstances (Mattys, Davis, Bradlow, \& Scott, 2012; Peelle, 2018).
The purpose of the current study is to explore novel word recognition by
adults under multimodal uncertainty. We focus on the special case where
people have access to multimodal cues from the auditory speech and the
visual referent.

One rigorous way to approach this question is through conducting an
\emph{ideal observer} analysis. This research strategy provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization.\footnote{It is, thus, a general
  instance of the rational approach to cognition (Anderson, 1990). It
  can also be seen as an instance of Marr's computational level of
  analysis (Marr, 1982).} When there is uncertainty in the input, the
ideal observer performs an optimal probabilistic inference. For example,
in order to recognize an ambiguous linguistic input, the model uses all
available probabilistic knowledge in order to maximize the accuracy of
this recognition. The ideal observer model can be seen as a theoretical
upper limit on performance. It is not so much a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003; Rahnev \& Denison, 2018). When there is a
deviation from the ideal, it can reveal extra constraints on human
cognition, such as limitations on the working memory or attentional
resources. The ideal observer analysis has had a tremendous impact not
only on speech related research (Clayards, Tanenhaus, Aslin, \& Jacobs,
2008; Feldman, Griffiths, \& Morgan, 2009; Kleinschmidt \& Jaeger, 2015;
Norris \& McQueen, 2008), but also on many other disciplines in the
cognitive sciences (for reviews, see Chater \& Manning, 2006; Knill \&
Pouget, 2004; Tenenbaum, Kemp, Griffiths, \& Goodman, 2011)

Some of these ideal-observer-based studies are closely related to the
question we are addressing in the current work. For instance, Clayards
et al. (2008) simulated auditory uncertainty by manipulating the
probability distribution of a cue (Voice Onset Time) that differentiated
similar words (e.g., \enquote{beach} and \enquote{peach}). They found
that humans were sensitive to these probabilistic cues and their
judgments closely reflected the optimal predictions. In another work,
Feldman et al. (2009) studied the perceptual magnet effect, a phenomenon
that involves reduced discriminability near prototypical sounds in the
native language (Kuhl, 1991). They showed that this effect can be
explained as the consequence of optimally solving the problem of
perception under uncertainty.

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the behavior of the ideal observer. This
previous research, however, did not systematically study speech
understanding when the visual information is obtained, not through the
speaker's facial features---as in audio-visual speech perception, but
through the referential context. In fact, experimental findings showed
that information about the identity of the semantic referent can be
integrated with linguistic information to resolve lexical and syntactic
ambiguities in speech (e.g., Eberhard, Spivey-Knowlton, Sedivy, \&
Tanenhaus, 1995; Spivey, Tanenhaus, Eberhard, \& Sedivy, 2002;
Tanenhaus, Spivey-Knowlton, Eberhard, \& Sedivy, 1995). To our
knowledge, however, no study offered an ideal observer analysis of word
identification in such context, that is, when the listener has to
combine cues from the sound and the referent.

On the face of it, the question of combining information from the sound
and the visual referent might seem similar to that of audio-visual
speech integration. Nevertheless, there are at least two fundamental
differences between these two cases, and both can influence the way the
auditory and visual cues are combined: First, in the case of
audio-visual speech, both modalities offer information about the same
underlying speech category. They may differ only in terms of their
informational reliability. In a referential context, however, the
auditory and visual modalities play different roles in the referential
process---in addition to possible differences in informational
reliability. Indeed, the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning}. It has been
shown that speech has a privileged status compared to other sensory
stimuli (Edmiston \& Lupyan, 2015; Lupyan \& Thompson-Schill, 2012;
Vouloumanos \& Waxman, 2014; Waxman \& Gelman, 2009; Waxman \& Markow,
1995), and that this privilege is indeed related to the speech's ability
to refer (Waxman \& Gelman, 2009).\footnote{There is, however, a debate
  as to whether speech is privileged for children and adults for similar
  reasons. Whereas some researchers suggest that speech is privileged
  for both children and adults because of its ability to refer (e.g.,
  Waxman \& Gelman, 2009), others suggest that speech might \emph{not}
  have a referential status from the start. Rather, speech might be
  preferred by children only because of a low level auditory
  ``overshadowing'' (e.g., Sloutsky \& Napolitano, 2003).} Thus, in a
referential context, it is possible that listeners do not treat the
auditory and visual modalities as equivalent sources of information.
Instead, there could be a sub-optimal bias for the auditory modality
beyond what is expected from informational reliability alone.

Second, in the case of audio-visual speech, the auditory and visual
stimuli are expected to be perceptually correlated. The expectation for
this correlation is such that when there is a mismatch between the
auditory and visual input, people still integrate them into a unified
(but illusory) percept (McGurk \& MacDonald, 1976). In the case of
referential language, however, the multimodal association is by nature
\emph{arbitrary} (Greenberg, 1957; Saussure, 1916). For instance, there
is no logical/perceptual connection between the sound \enquote{bee} and
the corresponding insect. Moreover, variation in the way the sound
\enquote{bee} is pronounced is generally not expected to correlate
perceptually with variation in the shape (or any other visual property)
in the category of bees. In sum, cue combination in the case of
arbitrary audio-visual associations (word-referent) is likely to be less
automatic, more effortful, and therefore less conducive to optimal
integration than it is in the case of perceptually correlated
associations (as in audio-visual speech perception).

@@

We investigate how people combine cues from the auditory and the visual
modality to recognize words in a referential context. In particular, we
study how this combination is performed under various degrees of
uncertainty in both the auditory and the visual modality. Imagine, for
example, that someone is uncertain whether they heard \enquote{pea} or
\enquote{bee}, does this uncertainty make them rely more on the referent
(e.g., the object being pointed at)? Vice versa, if they are not sure if
they saw a bee or a fly, does it make them rely more on the sound? More
importantly, when input in both modalities is uncertain to varying
degrees, do they weight each modality according to its relative
reliability, which is the optimal strategy, or do they over-rely on a
particular modality, which is a sub-optimal strategy?

We perform a rational analysis of the task. First we propose an ideal
observer model that performs the combination in an optimal fashion.
Second we compare the predictions of the optimal model to human
responses. Humans can deviate from the ideal for several reasons. For
instance, as mentioned above, a sub-optimality can be induced by the
suggested privileged status of speech or by the arbitrariness of the
referential association. In order to study possible patterns of
sub-optimality, we compare the optimal model (which provides a normative
benchmark) to a descriptive model (which is fit to human responses).
Comparing parameter estimates between these two formulations allows us
to quantify the degree of deviation from optimality.

We tested the ideal observer model's predictions in three behavioral
experiments where we varied the source of uncertainty. In Experiment 1,
audio-visual tokens were ambiguous with respect to their category
membership only. In Experiment 2, we intervened by adding background
noise to the auditory modality, and in Experiment 3, we intervened by
adding background noise to the visual modality. In all experiments,
participants were quantitatively near-optimal, though overall response
precision was slightly lower than expected. Moreover, in Experiment 1
where neither of the modalities was perturbed with background noise,
participants weighted auditory and visual cues according to the relative
reliability predicted by the optimal model. In other words, we found no
evidence for a modality bias towards either the auditory or the visual
modality. However, in Experiment 2 and 3, participants over-relied on
one modality when the other modality was perturbed with additional
noise.

@@

We introduce a task adapted from Sloutsky and Napolitano (2003). The
original task has been used with both children and adults to probe
audio-visual encoding (see Robinson \& Sloutsky, 2010 for a review).
Here we use a slightly different version to test word recognition in a
referential context. We use two visual categories (cat and dog) and two
auditory categories (/b/ and /d/ embedded in the minimal pair
/aba/-/ada/). For each participant, an arbitrary pairing is set between
the auditory and the visual categories, leading to two audio-visual word
categories (e.g., dog-/aba/, cat-/ada/). In each trial, participants are
presented with an audio-visual target (the prototype of the target
category), immediately followed by an audio-visual test stimulus
(Figure~\ref{fig:task}). The test stimulus may differ from the target in
both the auditory and the visual components. After these two
presentations, participants press \enquote{same} or \enquote{different.}

@@

In the testing phase of the original task (Sloutsky \& Napolitano,
2003), participants are asked whether or not the two audio-visual
presentations are \emph{identical}. In the current study, we are
interested, rather, in the categorization, i.e., determining whether or
not two similar tokens are members of the same phonological/semantic
category. Therefore, testing in our task is category-based: Participants
are asked to press \enquote{same} if they think the second item (the
test) belongs to the same category as the first (target) (e.g.,
dog-/aba/), even if there is a slight difference in the sound, in the
referent, or in both. They are instructed to press \enquote{different}
only if they think that the second stimulus was an instance of the other
category (cat-/ada/). The task also includes trials where pictures are
hidden (audio-only) or where sounds are muted (visual-only). These
unimodal trials provide us with the participants' evaluation of the
probabilistic information present in the auditory and visual categories.
As we shall see, these unimodal distributions are used as inputs to the
optimal cue combination model.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002)---which typically studies continuous stimuli (e.g., size,
location)---the probabilistic information in our case cannot be
characterized with \emph{sensory noise}, only. Indeed, our task involves
responses over categorical variables (phonemes and concepts), and
therefore, the optimal model should take into account, not only the
noise variability around an individual perceptual estimate, but also its
\emph{categorical variability}, i.e., the uncertainty related to whether
this perceptual estimate belongs to a given category (see also
Bankieris, Bejjanki, \& Aslin, 2017; Bejjanki et al., 2011). In what
follows, we describe a model that accounts for both types of
variability. First, we describe the model in the simplified case of
categorical variability only. Second, we augment this simplified model
to account for sensory noise.

@@

So far, we only accounted for categorical variability. For instance, if
the speaker generates a target production \(a_t\) from an auditory
category \(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})\), the ideal model
assumes that it has direct access to this production token (i.e.,
\(a=a_t\)), and that all uncertainty is about the category membership of
this token. However, we might also want to account for internal noise in
the brain and/or external noise in the environment. For example, the
observer might not have access to the exact produced target, but only to
the target perturbed by noise. If we assume this noise to be normally
distributed, that is, \(p(a | a_t) \sim N(a_t, \sigma^2_{N_A})\), then
integrating over \(a_t\) leads to this new expression of the probability
distribution: \[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})\]
Similarly, in the case of sensory noise in the visual modality, we get:
\[ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})\] Finally,
using exactly the same derivation as above, we end up with the following
multimodal weighting scheme in the optimal combination model (Equation
1) which takes into account both categorical and sensory variability:

@@

Equation 1 provides the optimal model's predictions for how
probabilities that characterize uncertainty in the auditory and the
visual modalities can be combined to make categorical decisions.
Parameter estimates of the probability distributions in each modality
are derived by fitting unimodal posteriors to the participants'
responses in the unimodal conditions, i.e., the condition where only the
sounds are heard or only the pictures are seen
(Figure~\ref{fig:task}).\footnote{Further technical detail about model fitting in the unimodal conditions will be given in the method section of Experiment 1}
Using these derived parameters, the optimal model makes predictions
about responses in the bimodal (i.e., audio-visual) condition where
participants both hear the sounds and see the pictures.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Both variances may increase, but their proportion remains the same.
  That is, \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), but
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} \approx \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimality would be due to increased randomness in
  human responses in the bimodal condition. However, this randomness
  would not affect the relative weighting of both modalities, i.e.,
  participants would still weigh modalities according to the relative
  reliability predicted by the optimal model.
\item
  The auditory variance increases at a higher rate. That is,
  \(\sigma^2_{Ab} \gg \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimally would consist not only in participants
  being more random in the bimodal condition, but also in having a
  systematic preference for the visual modality, even after accounting
  for informational reliability.
\item
  The visual variance increases at a higher rate. That is,
  \(\sigma^2_{Vb} \gg \sigma^2_{V}\), and
  \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  This case is the reverse of case 2, i.e., in addition to increased
  randomness in the bimodal condition, there is a systematic preference
  for the auditory modality, even after accounting for informational
  reliability.
\end{enumerate}

@@

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. They were paid at an hourly rate
of \$6/hour. Participants were excluded if they reported having
experienced a technical problem of any sort during the online experiment
(N=14), or if they had less than 50\% accurate responses on the
unambiguous training trials (N=6). The final sample consisted of N = 80
participants.\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

@@

@@

This work was supported by a post-doctoral grant from the Fyssen
Foundation + XXX

@@

\hypertarget{ref-dautriche2015}{}
Dautriche, I., Swingley, D., \& Christophe, A. (2015). Learning novel
phonological neighbors: Syntactic category matters. \emph{Cognition},
\emph{143}, 77--86.

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={How Optimal is Word-Referent Identification Under Multimodal Uncertainty?},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{How Optimal is Word-Referent Identification Under Multimodal
Uncertainty?}

@@

  \abstract{Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We introduce
a paradigm that allows us to examine how adults identify words under
joint uncertainty in the auditory and visual modalities and propose an
ideal observer model of how cues in these modalities are combined
optimally. Model predictions are tested in three experiments where word
recognition is made under two kinds of uncertainty: category ambiguity
and perceptual noise. In all cases, the optimal model explains much of
the variance in human mean judgments. When the signal is not distorted
with noise, participants weight the auditory and visual cues optimally,
that is, according to the relative reliability of each modality. But
when one modality has noise added to it, human perceivers systematically
prefer the unperturbed modality to a greater extent than the optimal
model does. The study provides a formal framework which helps to
quantify how word form and word meaning interact in word recognition
under uncertainty. Moreover it offers a first step towards a model that
accounts for form-meaning synergies in early word learning.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

@@

\section{Introduction}\label{introduction}

Language uses symbols expressed in one modality -- the auditory
modality, in the case of speech -- to communicate about the world, which
we perceive through many different sensory modalities. Consider hearing
someone yell \enquote{bee!} at a picnic, as a honey bee buzzes around
the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence. However, word identification
takes place in a noisy world, and the cues received through each
modality may not provide a definitive answer. On the auditory side,
individual acoustic word tokens are almost always ambiguous with respect
to the particular sequence of phonemes they represent, which is due to
the inherent variability of how a phonetic category is realized
acoustically (Hillenbrand, Getty, Clark, \& Wheeler, 1995). And some
tokens may be distorted additionally by mispronunciation or ambient
noise. Perhaps the speaker was yelling \enquote{pea} and not
\enquote{bee.} Similarly, a sensory impression may not be enough to make
a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

As a simplified case study of early word learning, the task of matching
sounds to corresponding visual objects has been studied extensively in
the developmental literature. For example, many studies focus on how
children might succeed in this type of task despite referential
ambiguity (Medina, Snedeker, Trueswell, \& Gleitman, 2011; Pinker, 1989;
Smith \& Yu, 2008; Suanda, Mugwanya, \& Namy, 2014; Vlach \& Johnson,
2013; Vouloumanos, 2008; Yurovsky \& Frank, 2015). However, even when
they \emph{know} the meanings of a word, listeners (both children and
adults) often still find it challenging to recognize which word the
speaker has uttered, especially under noise (Mattys, Davis, Bradlow, \&
Scott, 2012; Peelle, 2018). The purpose of the current study is thus to
explore word recognition by adults under multimodal uncertainty. We
focus on the special case where people have access to multimodal cues
from the auditory speech and the visual referent. In the General
Discussion, we return to the question of how these findings relate to
questions about word learning.

One rigorous way to approach this question is through conducting an
\emph{ideal observer} analysis. This research strategy provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization.\footnote{It is, thus, a general
  instance of the rational approach to cognition (Anderson, 1990),
  instantiating Marr's computational level of analysis (Marr, 1982).}
When there is uncertainty in the input, the ideal observer performs an
optimal probabilistic inference. For example, in order to recognize an
ambiguous linguistic input, the model uses all available probabilistic
knowledge in order to maximize the accuracy of this recognition. The
ideal observer model can be seen as a theoretical upper limit on
performance. It is not so much a realistic model of human performance,
as much as a baseline against which human performance can be compared
(Geisler, 2003; Rahnev \& Denison, 2018). When there is a deviation from
the ideal, it can reveal extra constraints on human cognition, such as
limitations on the working memory or attentional resources. This
approach has had a tremendous impact not only on speech-related research
(Clayards, Tanenhaus, Aslin, \& Jacobs, 2008; Feldman, Griffiths, \&
Morgan, 2009; Kleinschmidt \& Jaeger, 2015; Norris \& McQueen, 2008),
but also on many other disciplines in the cognitive sciences (for
reviews, see Chater \& Manning, 2006; Knill \& Pouget, 2004; Tenenbaum,
Kemp, Griffiths, \& Goodman, 2011)

Some prior ideal observer studies are closely related to the question we
are addressing in the current work. For instance, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (Voice Onset Time) that differentiated similar
words (e.g., \enquote{beach} and \enquote{peach}). They found that
humans were sensitive to these probabilistic cues and their judgments
closely reflected the optimal predictions. And Feldman et al. (2009)
studied the perceptual magnet effect, a phenomenon that involves reduced
discriminability near prototypical sounds in the native language (Kuhl,
1991), showing that this effect can be explained as the consequence of
optimally solving the problem of perception under uncertainty.

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the predictions of an ideal observer. This
previous research did not, however, systematically study speech
understanding when visual information was obtained through the
referential context rather than through observation of speaker's face.
Although some experimental findings show that information about the
identity of a referent can be integrated with linguistic information to
resolve lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995), to our knowledge no study has offered an ideal observer
analysis of this task.

Combining information between words and visual referents might seem
similar to audio-visual speech integration, but there are at least two
fundamental differences between these two cases, and both can influence
the way the auditory and visual cues are combined.

First, in the case of audio-visual speech, both modalities offer
information about the same underlying speech category. They differ only
in terms of their informational reliability. In a referential context,
however, the auditory and visual modalities play different roles in the
referential process -- the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning} (and these
differences are in addition to possible differences in informational
reliability). Further, speech is claimed to have a privileged status
compared to other sensory stimuli (Edmiston \& Lupyan, 2015; Lupyan \&
Thompson-Schill, 2012; Vouloumanos \& Waxman, 2014; Waxman \& Gelman,
2009; Waxman \& Markow, 1995), and that this privilege is suggested to
be specifically related to the ability to refer (Waxman \& Gelman,
2009).\footnote{There is, however, a debate as to whether speech is
  privileged for children and adults for similar reasons. Whereas some
  researchers suggest that speech is privileged for both children and
  adults because of its ability to refer (e.g., Waxman \& Gelman, 2009),
  others suggest that speech might \emph{not} have a referential status
  from the start. Rather, speech might be preferred by children only
  because of a low level auditory ``overshadowing'' (e.g., Sloutsky \&
  Napolitano, 2003).} Thus, in a referential context, it is possible
that listeners do not treat the auditory and visual modalities as
equivalent sources of information. Instead, there could be a sub-optimal
bias for the auditory modality beyond what is expected from
informational reliability alone.

Second, in the case of audio-visual speech, the auditory and visual
stimuli are expected to be perceptually correlated. The expectation for
this correlation is strong enough that when there is a mismatch between
the auditory and visual input, they are still integrated into a unified
(but illusory) percept (e.g., the McGurk Effect; McGurk \& MacDonald,
1976). In the case of referential language, however, the multimodal
association is by nature \emph{arbitrary} (Greenberg, 1957; Saussure,
1916). For instance, there is no logical or perceptual connection
between the sound \enquote{bee} and the corresponding insect. Moreover,
variation in the way the sound \enquote{bee} is pronounced is generally
not expected to correlate perceptually with variation in the shape (or
any other visual property) in the category of bees. In sum, cue
combination in the case of arbitrary audio-visual associations
(word-referent) is likely to be less automatic, more effortful, and
therefore less conducive to optimal integration than it is in the case
of perceptually correlated associations (as in audio-visual speech
perception).

@@

We investigate how cues from the auditory and the visual modality are
combined in recognizing words in a referential context. In particular,
we study how this combination is performed under various degrees of
uncertainty in both the auditory and the visual modality. Imagine, for
example, that someone is uncertain whether they heard \enquote{pea} or
\enquote{bee.} Does this uncertainty make them rely more on the referent
(e.g., the object being pointed at)? Or, if they are not sure if they
saw a bee or a fly, does this uncertainty make them rely more on the
sound? More importantly, when input in both modalities is uncertain to
varying degrees, do they weight each modality according to its relative
reliability (the optimal strategy), or do they over-rely on a particular
modality?

We begin by proposing an ideal observer model that performs the
combination in an optimal fashion. We then compare the predictions of
the optimal model to human responses. Humans can deviate from the ideal
for several reasons. For instance, as mentioned above, a sub-optimality
can be induced by the privileged status of a particular modality or by
the arbitrariness of the referential association. In order to study
possible patterns of sub-optimality, we compare the optimal model (which
provides a normative benchmark) to a descriptive model (which is fit to
human responses). Comparing parameter estimates between these two
formulations allows us to quantify the degree of deviation from
optimality.

We tested the ideal observer model's predictions in three behavioral
experiments where we varied the source of uncertainty. In Experiment 1,
audio-visual tokens were ambiguous with respect to their category
membership only. In Experiment 2, we intervened by adding perceptual
noise to the auditory modality, and in Experiment 3, we intervened by
adding perceptual noise to the visual modality. In all experiments,
participants were quantitatively near-optimal, though overall response
precision was slightly lower than expected. In Experiment 1 -- where
neither of the modalities was perturbed with background noise --
participants weighted auditory and visual cues according to the relative
reliability predicted by the optimal model. However, in Experiment 2 and
3, participants over-relied on one modality when the other modality was
perturbed with additional noise.

@@

We introduce a paradigm adapted from a task used by Sloutsky and
Napolitano (2003). The original was used with both children and adults
to probe audio-visual encoding (see Robinson \& Sloutsky, 2010 for
review). Here we use a slightly different version to test word
recognition in a referential context. We use two visual categories (cat
and dog) and two auditory categories (/b/ and /d/ embedded in the
minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is
set between the auditory and the visual categories, leading to two
audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each
trial, participants are presented with an audio-visual target (the
prototype of the target category), immediately followed by an
audio-visual test stimulus (Figure~\ref{fig:task}). The test stimulus
may differ from the target in both the auditory and the visual
components. After these two presentations, participants press
\enquote{same} or \enquote{different.}

@@

In the testing phase of the original task (Sloutsky \& Napolitano,
2003), participants were asked whether or not the two audio-visual
presentations are \emph{identical}. In the current study, we are
interested, rather, in the categorization, i.e., determining whether or
not two similar tokens are members of the same phonological/semantic
category. Therefore, testing in our task is category-based: Participants
are asked to press \enquote{same} if they think the second item (the
test) belongs to the same category as the first (target) (e.g.,
dog-/aba/), even if there is a slight difference in the sound, in the
referent, or in both. They are instructed to press \enquote{different}
only if they think that the second stimulus was an instance of the other
category (cat-/ada/). The task also includes trials where pictures are
hidden (audio-only) or where sounds are muted (visual-only). These
unimodal trials provide us with the participants' evaluation of the
probabilistic information present in the auditory and visual categories.
As we shall see, these unimodal distributions are used as inputs to the
optimal cue combination model.

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002), which typically studies continuous stimuli (e.g., size,
location), the probabilistic information in our case cannot be
characterized with \emph{sensory noise} only. Since our task involves
responses over categorical variables (phonemes and concepts), the
optimal model should take into account not only the noise variability
around an individual perceptual estimate but also its \emph{categorical
variability}, i.e., the uncertainty related to whether this perceptual
estimate belongs to a given category (see also Bankieris, Bejjanki, \&
Aslin, 2017; Bejjanki et al., 2011). In what follows, we describe a
model that accounts for both types of variability. First, we describe
the model in the simplified case of categorical variability only.
Second, we augment this simplified model to account for sensory noise.

@@

So far, we have only accounted for categorical variability. For
instance, if the speaker generates a target production \(a_t\) from an
auditory category \(p(a_t | A) \sim N(\mu_{A}, \sigma^2_{A})\), the
ideal model assumes that it has direct access to this production token
(i.e., \(a=a_t\)), and that all uncertainty is about the category
membership of this token. However, we might also want to account for
internal noise in the brain and/or external noise in the environment.
For example, the observer might not have access to the exact produced
target, but only to the target perturbed by noise. If we assume this
noise to be normally distributed, that is,
\(p(a | a_t) \sim N(a_t, \sigma^2_{N_A})\), then integrating over
\(a_t\) leads to this new expression of the probability distribution:
\[ p(a | A) \sim N(\mu_{A}, \sigma^2_{A}+\sigma^2_{N_A})\] Similarly, in
the case of sensory noise in the visual modality, we get:
\[ p(a | V) \sim N(\mu_{V}, \sigma^2_{V}+\sigma^2_{N_V})\] Finally,
using exactly the same derivation as above, we end up with the following
multimodal weighting scheme in the optimal combination model (Equation
1) which takes into account both categorical and sensory variability:

@@

Equation 1 provides the optimal model's predictions for how
probabilities that characterize uncertainty in the auditory and the
visual modalities can be combined to make categorical decisions.
Parameter estimates of the probability distributions in each modality
are derived by fitting unimodal posteriors to the participants'
responses in the unimodal conditions, i.e., the condition where only the
sounds are heard or only the pictures are seen
(Figure~\ref{fig:task}).\footnote{Further technical detail about model fitting in the unimodal conditions will be given in the method section of Experiment 1.}
Using these derived parameters, the optimal model makes predictions
about responses in the bimodal (i.e., audio-visual) condition where
participants both hear the sounds and see the pictures.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Both variances may increase, but their proportion remains the same.
  That is, \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), but
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} \approx \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimality would be due to increased randomness in
  human responses in the bimodal condition. However, this randomness
  would not affect the relative weighting of both modalities, i.e.,
  participants would still weigh modalities according to the relative
  reliability predicted by the optimal model.
\item
  The auditory variance increases at a higher rate. That is,
  \(\sigma^2_{Ab} \gg \sigma^2_{A}\) and
  \(\sigma^2_{Vb} \geqslant \sigma^2_{V}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  In this case, sub-optimality would consist not only in participants
  being more random in the bimodal condition, but also in having a
  systematic preference for the visual modality, even after accounting
  for informational reliability.
\item
  The visual variance increases at a higher rate. That is,
  \(\sigma^2_{Vb} \gg \sigma^2_{V}\), and
  \(\sigma^2_{Ab} \geqslant \sigma^2_{A}\), leading to
  \(\frac{\sigma^2_{Ab}}{\sigma^2_{Vb}} > \frac{\sigma^2_{A}}{\sigma^2_{V}}\).
  This case is the reverse of case 2, i.e., in addition to increased
  randomness in the bimodal condition, there is a systematic preference
  for the auditory modality, even after accounting for informational
  reliability.
\end{enumerate}

@@

We recruited a planned sample of 100 participants from Amazon Mechanical
Turk. Only participants with US IP addresses and a task approval rate
above 85\% were allowed to participate. They were paid at an hourly rate
of \$6/hour. Participants were excluded if they reported having
experienced a technical problem of any sort during the online experiment
(N=14), or if they had less than 50\% accurate responses on the
unambiguous training trials (N=6). The final sample consisted of N = 80
participants. All participants provided informed consent before taking
the experiment.
\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

@@

This work was supported by a post-doctoral grant from the Fyssen
Foundation.

\section{Disclosure statement}\label{disclosure-statement}

None of the authors have any financial interest or a conflict of
interest regarding this work and this submission.

@@

@@
